---
title: "Bike Share Project"
author: "Masoud A. Rostami"
date: "5/3/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is a project on Bike Share data in R-programming for data science nano-degree in Udaacity

```{r datasets}
ny = read.csv('new_york_city.csv')
wash = read.csv('washington.csv')
chi = read.csv('chicago.csv')
```

```{r library}
library(ggplot2)
library(pastecs)
library(tidyverse)
library(gridExtra)
```

Discriptive summary
```{r Discriptive summary}
str(ny)
str(wash)
str(chi)
summary(ny)
summary(wash)
summary(chi)
```



# Question 1: 
## Is there any different in trip duration between female and male in New York and Chicago? and how we can compare them together? 

Exploring the data
```{r Exploring the data}
objects(ny) 
objects(chi) 
ny1=na.omit(ny)
chi1=na.omit(chi)
```

Duration of trip by Gender in New York and Chicago
```{r Duration of trip by Gender in New York and Chicago}
by(ny1$Trip.Duration, ny1$Gender, summary) 
by(chi1$Trip.Duration, chi1$Gender, summary) 
```


Histogram to find the relationshib between the Trip duration and gender in New York and Chicago
```{r Histogram to find the relationshib between the Trip duration and gender in New York and Chicago}
Ny.f=subset(ny1, Gender=="Female")
Ny.m=subset(ny1, Gender=="Male")
chi.f=subset(chi1, Gender=="Female")
chi.m=subset(chi1, Gender=="Male")

par(mfrow = c(2, 2))  

hist(Ny.f$Trip.Duration,border="brown", breaks=5000, xlim=c(0,5000),  main="Trip duration in New York in Females", xlab="Trip duration")
hist(Ny.m$Trip.Duration, border="gray", breaks=50000, xlim=c(0,5000),  main="Trip duration in New York in Males", xlab="Trip duration")
hist(chi.f$Trip.Duration,border="blue", breaks=5000, xlim=c(0,5000),  main="Trip duration in Chicago in Females", xlab="Trip duration")
hist(chi.m$Trip.Duration, border="orange", breaks=5000, xlim=c(0,5000),  main="Trip duration in Chicago  in Male", xlab="Trip duration")
```

Boxplot  in trip duration by different genders in Chicago
```{r Boxplot  in trip duration by different genders in Chicago}
ggplot(data=subset(ny, !is.na(Gender)), aes(x=Gender, y=Trip.Duration))+geom_boxplot(aes(color=Gender))+ 
        coord_cartesian(ylim = c(0,2000))+ggtitle("Trip duraion by Gender in New york")+labs(y="Trip duration", x = "Gender")

ggplot(data=subset(chi, !is.na(Gender)), aes(x=Gender, y=Trip.Duration))+geom_boxplot(aes(color=Gender))+ 
        coord_cartesian(ylim = c(0,2000))+ggtitle("Trip duraion by Gender in Chicago")+labs(y="Trip duration", x = "Gender")
```



### Conclusion:
It seems in New York and Chicago, female are more interested to biking in Bike share company, 
and in the New York it is more common than Chicago. Female has the highest mean and median compare to the males.
Also, our histgram and boxplot sshow in both states the bike sharing is most popular in Female than men. 
it seems females are enjoying biking and having more fun.



# Questions 2:
## Which ages has the highest trip duration by using bike sharing? compare in New York and Chicago


Create a function called phist to unserstand the age of customer using Bike share
```{r Create a function called phist to unserstand the age of customer using Bike share}
phist <- function(x, ...) {
        hist(x,
             col = gray(.5, .2),
             border = "red",
             yaxt = "n",
             ylab = " ",
             ...)
        ci <- t.test(x)$conf.int
        top.text <- paste(
                "Mean = ", round(mean(x), 1),
                " (95% CI [", round(ci[1], 1),
                ", ", round(ci[2], 1), 
                "]), SD = ", round(sd(x), 1), 
                sep = " ")
        mtext(top.text, side = 3)
}
par(mfrow=c(2,1))
 
phist(ny1$Birth.Year, xlab = "Birth of the Year", main = "The Age of customers in New York", breaks=100)
phist(chi1$Birth.Year, xlab = "Birth of the year", main = "The Age of customers in Chicago", breaks=100)
```

Scatterplot
```{r Scatterplot}
ggplot(data=ny1, aes(x=Birth.Year, y=Trip.Duration))+geom_point(alpha=0.05, position = position_jitter(h=0)) +
        coord_trans(y='sqrt') +geom_jitter(stat = 'summary',color="red", fun = "median")+ylim(0,5000)+xlim(1939, 2000)+
        ggtitle("Trip duraion by Ages in New York")+labs(y="Trip duration", x = "Birth of the Year")



ggplot(data=chi1, aes(x=Birth.Year, y=Trip.Duration))+geom_point(alpha=0.05, position = position_jitter(h=0)) +
        coord_trans(y='sqrt') +geom_jitter(stat = 'summary',color="red", fun = "median")+ylim(0,5000)+xlim(1939, 2000)+
        ggtitle("Trip duraion by Ages in New York")+labs(y="Trip duration", x = "Birth of the Year")

```

Linear regression model1
```{r  Linear regression model1}
Model1=lm(Trip.Duration~Birth.Year, data=chi)
summary(Model1)


Model2=lm(Trip.Duration~Birth.Year, data=ny)
summary(Model2)
```



###  Conclusion:
It seems thae average age of birth of individuals using Bike share in New York is 1978 (Average 22 Years old) and in Chicago
is 1981 (Average 19 years old) that show more younger people in Chicago are interested inn using Bike.
Also, I did Linear model by using lm function and based on that both have 
high p-value that show we should ratin our null hypothesis mean there is not significant difference
between trip duration in different years. However, the p-value in chicaco is lower compare to
New york that may show in chicago may age have higher effect on trip duraion compare to New York.






# Question3: 
## Is there any different between user type and their trip duration between New York and Chicago?

Subsetting
```{r  Subsetting}
customer_ny=subset(ny1, User.Type=="Customer") #Subset the data for customer user type in New York
Subscriber_ny=subset(ny1, User.Type=="Subscriber") #Subset the data for Subscriber user type in New York
customer_chi=subset(chi1, User.Type=="Customer", na.rm=TRUE ) #Subset the data for customer user type in Chicago
Subscriber_chi=subset(chi1, User.Type=="Subscriber") #Subset the data for Subscriber  user type in Chicago
```



Create the function mean_my
```{r Create the function mean_my}
mean_my <- function(x) {
        output<- sum(x) / length(x) 
        return(output) 
        
}


mean_my(customer_ny$Trip.Duration)
mean_my(Subscriber_ny$Trip.Duration)
mean_my(customer_chi$Trip.Duration)
mean_my(Subscriber_chi$Trip.Duration)
```

Histogram in trip duraion
```{r Histogram in trip duraion}
a=ggplot(data=ny, aes(x=Trip.Duration))+geom_histogram(binwidth=10, alpha=0.2, col="blue")+facet_wrap(~User.Type)+xlim(0, 4000)+
        ggtitle("Trip duraion by User type in New York")+labs(y="Trip duration", x = "User Type")

#T
b=ggplot(data=chi, aes(x=Trip.Duration))+geom_histogram(binwidth=10, alpha=0.2, col="red")+facet_wrap(~User.Type)+xlim(0, 4000)+
        ggtitle("Trip duraion by User type in Chicago")+labs(y="Trip duration", x = "User Type")

grid.arrange(a,b , ncol = 1, nrow = 2)
```

 Linear regression model2
```{r Linear regression model2}
model.ny=(lm(Trip.Duration~User.Type, data=ny))
summary(model.ny)

model.chi=(lm(Trip.Duration~User.Type, data=chi))
summary(model.chi)
```


### Conclusion
It seems people that subscribe with Bikeshare company in New Yorka and Chicago are 
using more bikes compare to people with no subscirption, and  have the highest trip dutaion time.
Also linear model show the huge significant differnt between subscribe and no subscribe
trip duration due to very small p-value (< 2.2e-16) and high F-value. Also, our graph show that individuals with subscription in New York
are more than Chicago.

















